Theorems and Definitions

### 2016-03-26 math_modeling/class/notes/stochastics.tex
polynomials are fun

\begin{quote}
\textit{\textbf{Example}}\\
Consider the movement of people from the city to the suburb.
\begin{center}
\begin{tabular}{c | c c}
& Cities & Suburbs\\
\hline
Cities & 0.96 & 0.04\\
Suburbs & 0.1 & 0.99
\end{tabular}
\end{center}
with the initial vector
\[
\bf{x}_{0}=[60,125].
\]
The current model assumes that population stabilizes, but this is not always the case. 
\end{quote}

\end{document}

>>>>

\begin{quote}
\textit{\textbf{Example}}\\
Consider the movement of people from the city to the suburb.
\begin{center}
\begin{tabular}{c | c c}
& Cities & Suburbs\\
\hline
Cities & 0.96 & 0.04\\
Suburbs & 0.1 & 0.99
\end{tabular}
\end{center}
with the initial vector
\[
\bf{x}_{0}=[60,125].
\]
We can introduce some ``noise" into the system by discussing the eigenvalues of this system. In the
usual case, $\lambda=1$. We can show this by solving the system.
\begin{align*}
(\bf{P}-\lambda)\bf{x}&\to
\left[
\begin{array}{c c c}
-0.4 & 0.4 & 0\\
0.1 & 0.1 & 0
\end{array}
\right]\\
&\to
\left[
\begin{array}{c c c}
1 & -1 & 0\\
0 & 0  & 0
\end{array}
\right],
\end{align*}
so $x=y=1$.
\end{quote}

Often in Markov processes, we come across matrices raised to a power. To help us calculate these
products, we can come up with a new matrix, $\bf{S}$. We can then calculate
$\bf{S}^{-1}\bf{D}\bf{S}$, where $D$ is diagonalized matrix of the eigenvalues of $\bf{P}$. We know
that we can diagonalize $\bf{P}$ by the following theorem:
\begin{quote}
An $n\times n$ matrix $\bf{P}$ is diagonalizable if and only if it has $n$ linearly independent
eigenvectors.  
\end{quote}
\end{document}

### 2016-03-26 number_theory/hw/hw7.tex
polynomials are fun

\item Find all solutions to the congruence $x^{1000}\equiv1\mod10$.
\begin{quote}
By the prime factorization of 10, we have that
\[
5\mid(x^{1000}-1)\qquad\qquad\text{and}\qquad\qquad2\mid(x^{1000}-1).
\]
Since $2\mid(x^{1000}-1)$, we have that $x^{1000}\equiv1\mod2$, implying that $x$ is odd. This leaves
us with two possibilities in $\ZZ_{5}$, namely $\{1,3\}$. We can express 1000 as $5\cdot5\cdot5\cdot8$,
so we have
$((((x)^{8})^{5})^{5})^{5}\equiv(((x)^{8})^{5})^{5}\equiv((x)^{8})^{5}\equiv x^{8}\equiv1\mod5$.
We can express 8 as $4\cdot2$, so we have $(x^{2})^{4}\equiv1\mod5$, which means $x^{2}\equiv1\mod5$
by Fermat's Little Theorem. We know $x$ is either 1 or 3, so we can test the values. If we try 3, we
get that $9\equiv1\mod5$, which is not true, whereas if we try 1, we get $1\equiv1\mod5$, which is true.
Therefore, $x=1$.
\end{quote}
\end{enumerate}
\end{document}

>>>>

\item Find all solutions to the congruence $x^{1000}\equiv1\mod10$.
\begin{quote}
By the prime factorization of 10, we have that
\[
5\mid(x^{1000}-1)\qquad\qquad\text{and}\qquad\qquad2\mid(x^{1000}-1).
\]
Since $2\mid(x^{1000}-1)$, we have that $x^{1000}\equiv1\mod2$, implying that $x$ is odd. This leaves
us with two possibilities in $\ZZ_{5}$, namely $\{1,3\}$. We can express 1000 as $5\cdot5\cdot5\cdot8$,
so we have
$((((x)^{8})^{5})^{5})^{5}\equiv(((x)^{8})^{5})^{5}\equiv((x)^{8})^{5}\equiv x^{8}\equiv1\mod5$.
We can express 8 as $4\cdot2$, so we have $(x^{2})^{4}\equiv1\mod5$, which means $x^{2}\equiv1\mod5$
by Fermat's Little Theorem. We know $x$ is either 1 or 3, so we can test the values. If we try 3, we
get that $9^{4}\equiv1\mod5$, so $(-1)^{4}\equiv1\mod5$ and $1\equiv1\mod5$. If we try 1, we
get $1\equiv1\mod5$, which is true. Therefore, $x=1$ and $x=3$ are solutions. We can represent all
solutions as $1+10n,\ n\in\ZZ$ and $3+10k,\ k\in\ZZ$.
% If we consider the two
% other odd options in $\ZZ_{10}$, we have that
% $(7^{2})^{4}\equiv((-3)^{2})^{4}\equiv9^{4}\equiv(-1)^{4}\equiv1\mod10$. Similarly, if we allow
\end{quote}
\end{enumerate}
\end{document}
